---
title: "Dengue Situation in Bangladesh"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    social: menu
    source_code: https://github.com/hossainlab/dengue-situation-dashboard
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
# dashboard 
library(flexdashboard)
library(tidyverse)
library(readxl)
library(sjPlot)
library(viridis)
library(rio)
library(viridis)
library(denguedatahub)
library(visdat)

# Maps and Plots
library(plotly)
library(leaflet)
library(leafpop)
library(RColorBrewer)

# Dates
library(lubridate) 

## For displaying tables
library(DT)
library(readxl)

theme_set(theme_classic())
```


```{r, colors, include=FALSE}
confirmed_color <- "#FF5733"
active_color <- "#FFD700"
recovered_color <- "#00FF00"
death_color <- "#333333"
cfr_color <- "red"
```

```{r, data}
# cases data 
dengue_cases <- read_excel(here::here("data/DengueCaseReporting[2012-23].xlsx"),sheet = 1)
dengue_cases <- dengue_cases |> mutate(Months = factor(Months, levels = month.name)) 

# deaths data 
dengue_deaths <- read_excel(here::here("data/DengueCaseReporting[2012-23].xlsx"),sheet = 2)
dengue_deaths <- dengue_deaths |> mutate(Months = factor(Months, levels = month.name)) 

# daily data for 2023 
daily_data_23 <- import_list(here::here("data/DengueCases2023.xlsx"),
                             setclass = "data.table", 
                             rbind_label = "file", 
                             rbind = TRUE)
# convert into date format 
daily_data_23$Date <- as.Date(daily_data_23$Date)
# impute missing values with zero 
daily_data_23[is.na(daily_data_23)] <- 0

# daily_data_23
daily_data_23 <- daily_data_23 |> 
  mutate(Month = lubridate::month(Date, label = TRUE),
         Day = lubridate::day(Date)) 

# transform data into long format for visualizations 
# cases 
long_cases_data <- dengue_cases |> 
  pivot_longer(cols = 2:13, names_to = "Year", values_to = "Cases") |> 
  mutate(Months = factor(Months, levels = month.name)) 

# deaths 
long_deaths_data <- dengue_deaths |> 
  pivot_longer(cols = 2:10, names_to = "Year", values_to = "Deaths") |> 
  mutate(Months = factor(Months, levels = month.name)) 


# join data 
all_years <- left_join(long_cases_data, long_deaths_data, 
                       by = c("Months", "Year"))
all_years$Year <- as.factor(all_years$Year)

# Monthly dengue cases for 2023 
long_cases_data_23 <- long_cases_data|> 
  filter(Year == 2023)

long_deaths_data_23 <- long_deaths_data|> 
  filter(Year == 2023)

# join data for monthly cases 
data23 <- left_join(long_cases_data_23, long_deaths_data_23, 
                    by = c("Months", "Year"))

data23$Year <- as.factor(data23$Year)
```



Outbreak 2023  
===================================================================

Row
-------------------------------------------------------------------
### Confirmed Cases {.value-box}

```{r}
confirmed <- format(sum(daily_data_23$Cases, na.rm = T), big.mark = ",")
valueBox(value = confirmed, 
         caption = "Total Confirmed Cases", 
         color = confirmed_color, 
         icon = "fa-virus")
```


### Confirmed Deaths {.value-box}
```{r}
deaths <- format(sum(daily_data_23$Deaths, na.rm = T), big.mark = ",")
valueBox(value = deaths, 
         caption = "Total Death Cases", 
         color = death_color, 
         icon = "fa-exclamation-triangle")
```

### Case Fatality Rate (CFR) {.value-box}
```{r}
cfr <-  round(sum(daily_data_23$Deaths, na.rm = T) / sum(daily_data_23$Cases, na.rm = T) * 100, 2)

valueBox(value = cfr, 
         caption = "Case Fatality Rate (CFR)", 
         color = cfr_color, 
         icon = "fa-heart-broken")
```


### Recovered Cases {.value-box}
```{r}
recovered <- format(sum(daily_data_23$Recovered, na.rm = T), big.mark = ",")
valueBox(value = recovered, 
         caption = "Total Recovered Cases", 
         color = recovered_color, 
         icon = "fa-check-circle")
```


### Active Cases {.value-box}
```{r}
daily_data_23$Active <- daily_data_23$Cases - daily_data_23$Deaths - daily_data_23$Recovered
active <- format(sum(daily_data_23$Active, na.rm = T), big.mark = ",")
valueBox(value = active, 
         caption = "Total Active Cases", 
         color = active_color, 
         icon = "fa-skull")
```


Row
-----------------------------------------------------------------------
### Daily Confirmed, Death, and Recovered Cases
```{r}
# Viz 1: Daily Confirmed, Death, and Recovered Cases Over Time in 2023
plot_ly(data = daily_data_23, x = ~Date, type = 'scatter', mode = 'lines+markers', 
        name = 'Cases', y = ~Cases) %>%
  add_trace(y = ~Deaths, name = 'Deaths', line = list(color="black")) %>%
  add_trace(y = ~Recovered, name = 'Recovered') %>%
  layout(title = "Daily Dengue Cases", 
         xaxis = list(title = 'Date'),
         yaxis = list(title = 'Cases'), 
         legend = list(orientation = "h", 
                       xanchor = "center", 
                       x = 0.5, 
                       y = -0.3))
```

### Monthly Dengue Cases, Deaths, and Recovered
```{r, include=FALSE}
monthly_data_23 <- daily_data_23 |> 
  group_by(Month) |> 
  summarise(TotalCase = sum(Cases), 
            TotalDeaths = sum(Deaths), 
            TotalRecovered = sum(Recovered))
```

```{r}
# Viz 2: Monthly Dengue Cases, Deaths, and Recovered 
plot_ly(data = monthly_data_23, x = ~Month, type = 'scatter', mode = 'lines+markers', 
        name = 'Cases', y = ~TotalCase) |> 
  add_trace(y = ~TotalDeaths, name = "Deaths", type = 'scatter', mode = 'lines+markers') |> 
  add_trace(y = ~TotalRecovered, name = "Recovered", type = 'scatter', mode = 'lines+markers') |> 
  layout(title = "Monthly Dengue Cases", 
         xaxis = list(title = 'Date'),
         yaxis = list(title = 'Cases'), 
         legend = list(orientation = "h", 
                       xanchor = "center", 
                       x = 0.5, 
                       y= -0.3))

```


Row
-------------------------------------------------------------------

### Number of Dengue Cases

```{r}
# Viz 3: Number of Dengue Cases
plot_ly(data = data23, x = ~Months)  |> 
  add_bars(y = ~Cases, name = 'Confirmed Cases') |> 
  add_trace(y = ~Cases, type = 'scatter', mode = 'lines+markers', name = "Trends") |> 
  layout(title = "Monthly Dengue Cases",
         xaxis = list(title = 'Month'),
         yaxis = list(title = 'Number of Confirmed Cases'), 
         legend = list(orientation = "h", 
                       xanchor = "center", 
                       x = 0.5, 
                       y= -0.3))
```


### Number of Death Cases
```{r}
# Viz 4: Number of Death Cases in 2023
plot_ly(data = data23, x = ~Months) |> 
  add_bars(y = ~Deaths, name = 'Confirmed Deaths', marker = list(color = 'black')) |> 
  add_trace(y = ~Deaths, type = 'scatter', mode = 'lines+markers', name = "Trends") |> 
  layout(title = "Monthly Dengue Deaths",
         xaxis = list(title = 'Month'),
         yaxis = list(title = 'Number of Death Cases'), 
         legend = list(orientation = "h", 
                       xanchor = "center", 
                       x = 0.5, 
                       y = -0.3))
```

All Outbreaks
===================================================================


Row
-----------------------------------------------------------------------
### Number of Confirmed Dengue Cases by Month and Year [2012-2023]
```{r, include=FALSE}
all_years_aggregated <- all_years |> 
  group_by(Year, Months) |> 
  summarise(TotalCase = sum(Cases), 
            TotalDeaths = sum(Deaths))
```

```{r}
# Viz 5: Monthly Total Cases by Month and Year [2012-2023]
plot_ly(data = all_years_aggregated, x = ~Months, y = ~TotalCase, type = 'scatter', 
        mode = 'lines', color = ~Year) |> 
  layout(title = "Monthly Total Cases by Year from 2012 to 2023",
         xaxis = list(title = "Month"),
         yaxis = list(title = "Total Cases"), 
         legend = list(title=list(text='<b> Years </b>'))) 
```

### Number of Dengue Death Cases by Month and Year [2015-2023]
```{r}
# Viz 6: Monthly Total Deaths by Year [2015-2023]
plot_ly(data = all_years_aggregated, x = ~Months, y = ~TotalDeaths, type = 'scatter', mode = 'lines', color = ~Year) |> 
  layout(title = "Monthly Total Deaths by Year from 2015 to 2023",
         xaxis = list(title = "Month"),
         yaxis = list(title = "Total Deaths"), 
         legend = list(title=list(text='<b> Years </b>'))) 
```

Row
-------------------------------------------------------------------

### Year-wise Dengue Cases[2012-2023]
```{r, include=FALSE}
year_wise_aggregated <- all_years |> 
  group_by(Year) |> 
  summarise(Total_Cases = sum(Cases), 
            Toal_Deaths = sum(Deaths))
year_wise_aggregated
```


```{r}
# Viz 7: Year-wise Dengue Cases in Bangladesh from 2012 to 2023
plot_ly(data = year_wise_aggregated, x = ~Year)  |> 
  add_bars(y = ~Total_Cases, name = 'Cases', marker = list(color = "#2f4b7c")) |> 
  layout(title = 'Year-wise Dengue Cases from 2012 to 2023', 
         xaxis = list(title = 'Year'),
         yaxis = list(title = 'Number of Deaths'))
```

### Year-wise Dengue Death Cases [2015-2023]
```{r}
# Viz 8: Year-wise Dengue Cases in Bangladesh from 2015 to 2023
plot_ly(data = year_wise_aggregated, x = ~Year)  |> 
  add_bars(y = ~Toal_Deaths, name = 'Cases', marker = list(color = "#ff7c43")) |> 
  layout(title = 'Year-wise Dengue Death Cases from 2015 to 2023', 
         xaxis = list(title = 'Year'),
         yaxis = list(title = 'Number of Cases'))

```



Top 5 - Outbreaks 
===================================================================

Row
-------------------------------------------------------------------
### Top 5 Dengue Outbreaks Cases by Year and Month 

```{r, include=FALSE}
# Top 5 outbreaks by month and year in Bangladesh 
top5_outbreaks_month_year <- all_years |> 
  group_by(Year, Months) |> 
  summarise(TotalCases = sum(Cases), 
            TotalDeaths = sum(Deaths)) |> 
  filter(Year %in% c(2018, 2019, 2021, 2022, 2023))
```

```{r, include=FALSE}
# Top 5 outbreaks by year in Bangladesh 
top5_outbreaks_year <- all_years |> 
  group_by(Year) |> 
  summarise(TotalCases = sum(Cases), 
            TotalDeaths = sum(Deaths)) |> 
  filter(Year %in% c(2018, 2019, 2021, 2022, 2023))
```


```{r}
# Top 5 Dengue Outbreaks by Year and Month 
plot_ly(data = top5_outbreaks_month_year, x = ~Months, y = ~TotalCases, type = 'scatter', mode = 'lines', color = ~Year) |> 
  layout(title = "Top 5 Dengue Outbreaks Cases by Year and Month",
         xaxis = list(title = "Month"),
         yaxis = list(title = "Total Cases"), 
         legend = list(title=list(text='<b> Years </b>')))
```


### Top 5 Dengue Outbreaks Cases by Year

```{r}
# Year-wise dengue cases 
plot_ly(data = top5_outbreaks_year, x = ~Year)  |> 
  add_bars(y = ~TotalCases, name = 'Cases', marker = list(color = "#5ec962")) |> 
  layout(title = 'Top 5 Dengue Outbreaks Cases by Year', 
         xaxis = list(title = 'Year'),
         yaxis = list(title = 'Number of Cases'))

```


Row
-------------------------------------------------------------------
### Top 5 Dengue Outbreaks Deaths by Month and Year

```{r}
plot_ly(data = top5_outbreaks_month_year, x = ~Months, y = ~TotalDeaths, type = 'scatter', mode = 'lines', color = ~Year) |> 
  layout(title = "Monthly Total Death Cases for top 5 Outbreaks",
         xaxis = list(title = "Month"),
         yaxis = list(title = "Total Cases"), 
         legend = list(title=list(text='<b> Years </b>')))
```


### Top 5 Dengue Outbreaks Deaths by Year

```{r}
# Year-wise dengue cases 
plot_ly(data = top5_outbreaks_year, x = ~Year)  |> 
  add_bars(y = ~TotalDeaths, name = 'Cases', marker = list(color = "#21918c")) |> 
  layout(title = 'Year-wise Dengue Deaths', 
         xaxis = list(title = 'Year'),
         yaxis = list(title = 'Number of Cases'))
```

Global Situation
===================================================================

Row
-------------------------------------------------------------------
### Level of Dengue Risk around the Globe 

```{r, include=FALSE}
tab1 <- level_of_risk %>% group_by(region, level_of_risk) %>%
  summarise(count = n(), .groups = 'drop') 
knitr::kable(tab1)

```

```{r}
# Summarize and Visualize level_of_riskdata
p <- ggplot(tab1, aes(region, count)) + 
  geom_bar(aes(fill = level_of_risk), stat = "identity", position = "dodge")+
  labs(title = "Level of Dengue risk around the world",
       subtitle = "The level of risk in the context of dengue refers to the likelihood of dengue outbreaks happening in various regions worldwide.", 
       x = "Region",
       y = "Number of countries", 
       caption = "by Jubayer Hossain, Data Source: denguedatahub")+ 
  guides(fill = guide_legend(title = "Level of Risk"))+
  scale_fill_brewer(palette = "Dark2")+ 
  theme(legend.position = "top") 
ggplotly(p)
```

### Presence of Dengue Incidence in 2019
```{r}
# Presence of dengue incidence in 2019
worlddata2019 <- dplyr::filter(world_annual, year == 2019)
p2 <- ggplot(worlddata2019, aes(x = long,
                          y = lat,
                          group=group,
                          fill = dengue.present)) +
  geom_polygon(color = "black") +
  labs(title = "Presence of dengue incidence in 2019", 
       subtitle = "Dengue Incidence in 2019: A Global Health Challenge", 
       x = "Longitude",
       y = "Latitude", 
       caption = "by Jubayer Hossain, Data Source: denguedatahub") +
  guides(fill = guide_legend(title = "Dengue Present?"))+
  theme(legend.position = "top") + 
  scale_fill_brewer(palette = "Dark2") 

ggplotly(p2)
```



Dengue - Climate Change 
===================================================================





Data
===================================================================
**Data**

The Dengue Situation Dashboard utilizes data from the official website of the [Directorate General of Health Services (DGHS)](https://old.dghs.gov.bd/index.php/bd/home/5200-daily-dengue-status-report) in Bangladesh, specifically the Daily Dengue Status Report. Developed using the flexdashboard framework, this dashboard provides valuable insights and real-time information about the dengue situation in Bangladesh.

The dashboard presents a user-friendly interface that allows users to access and visualize the latest data related to dengue cases, deaths, and affected areas in the country. By incorporating the data from the DGHS, the dashboard ensures the reliability and accuracy of the information presented.

Through interactive charts, graphs, and maps, the dashboard highlights key statistics, trends, and patterns related to dengue cases. Users can explore the data by region, district, or specific time periods, gaining a comprehensive understanding of the dengue situation at various levels.

The dashboard offers a range of visualizations, including line charts displaying the daily or weekly trend of dengue cases, bar charts showcasing the distribution of cases across different districts, and choropleth maps indicating the severity of dengue outbreaks in different regions of Bangladesh. These visual representations help stakeholders and decision-makers to identify high-risk areas, monitor the progression of the disease, and allocate resources effectively.


Furthermore, the Dengue Situation Dashboard allows users to compare current data with historical data, enabling them to assess the impact of control measures and interventions. This feature assists in evaluating the effectiveness of strategies implemented to mitigate the dengue outbreak.

By providing real-time updates and an intuitive interface, the Dengue Situation Dashboard based on the DGHS data ensures that stakeholders, including health officials, researchers, and the public, have access to reliable and up-to-date information. This empowers them to make informed decisions, implement targeted interventions, and raise awareness about dengue prevention and control measures.

Overall, the Dengue Situation Dashboard based on the DGHS data and developed using flexdashboard is a valuable tool that enhances the understanding of the dengue situation in Bangladesh. By leveraging the power of data visualization and real-time information, this dashboard supports evidence-based decision-making and facilitates effective response measures to combat dengue outbreaks in the country.


Dengue - Bangladesh
=======================================================================
**Outbreak at a glance** 

As of 20 November 2022, a total of 52 807 laboratory-confirmed dengue cases and 230 related deaths have been reported by the Ministry of Health & Family Welfare of Bangladesh since 1 January 2022 with a case fatality rate (CFR) of 0.44%. Dengue is endemic in Bangladesh however a surge of cases started in June 2022. Currently, all eight divisions in the country are reporting cases and deaths. This is the second-largest outbreak since 2000, with the largest having occurred in 2019. The current dengue outbreak is unusual in its scale and seasonality.

**Description of the outbreak**

Between 1 January and 20 November 2022, a total of 52 807 dengue cases including 230 related deaths (case fatality rate = 0.44%) were reported by the Ministry of Health & Family Welfare (MOHFW) (Figure 1). The cases were confirmed either by non-structural protein (NS1) diagnostic kits or by Immunoglobulin M (IgM) tests. According to information available for 40% of reported cases (n=20 982) the median age is 25 years (range: 0 – 89) with males accounting for 60% of the cases. This is the second highest annual number of cases since 2000, the highest having occurred in 2019, when 101 354 cases including 164 deaths were reported (Figure 2 & 3).

The most affected division is Dhaka, accounting for 70.6% of cases and 60.4% of deaths. Dhaka city, the largest city in Bangladesh, located in Dhaka division, has reported 64.5% (n= 34 071) of the total number of cases. Other affected divisions include Chattogram division (13.2% of cases and 24.8% of deaths) and Khulna division (5.5% of cases and 4.8% of deaths) (Figure 4).

The high incidence of dengue cases this year is taking place in the context of an unusual amount of rainfall since June 2022, accompanied by high temperatures and high humidity which have resulted in an increased mosquito population throughout Bangladesh.

Source: https://www.who.int/emergencies/disease-outbreak-news/item/2022-DON424



About  
=======================================================================
**Author**

[Md. Jubayer Hossain](https://hossainlab.github.io/) is the Founder, Management Lead and Data Analyst at CHIRAL Bangladesh. CHIRAL Bangladesh is a non-profit organization dedicated to health research to improve lives in Bangladesh. He aspires to maximize the quality of life of people around him by working at the intersection of education, technology, and health research. His research interests include public health, health data science, remote sensing, GIS applications in health,  and machine learning in healthcare. He has been teaching programming, data analysis, data science, and research methodology since 2020 in CHIRAL Bangladesh. Detailed research and teaching activities were found on his [website.](https://hossainlab.github.io/)  




